---
title: "Convergence & Stabilisation"
author: "Federica"
date: "2026-09-01"
format:
  html:
    embed-resources: true
    code-fold: true
    toc: true
    df-print: kable
  pdf: default
execute:
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(knitr)
library(patchwork)   # install.packages("patchwork") if missing — side-by-side plots
library(sf)          # install.packages("sf")        — course-context maps

theme_set(theme_minimal(base_size = 12))

outputs_dir <- "/Users/federicadomecq/gitrepos/golfOnPar/Parallelisation/convergence/outputs"

data_dir <- "/Users/federicadomecq/gitrepos/golfOnPar/Parallelisation/convergence/MyScripts/data/"

equiv_sequential   <- read_csv(file.path(data_dir, "equiv_sequential.csv"))
equiv_cross_seed   <- read_csv(file.path(data_dir, "equiv_cross_seed.csv"))
equiv_overall      <- read_csv(file.path(data_dir, "equiv_overall_summary.csv"))
equiv_nonconv      <- read_csv(file.path(data_dir, "equiv_non_converged_gridpoints.csv"))
stabilisation_N    <- read_csv(file.path(data_dir, "stabilisation_vs_N.csv"))
stabilisation_seed <- read_csv(file.path(data_dir, "stabilisation_all_seeds.csv"))
cross_seed_summary <- read_csv(file.path(data_dir, "cross_seed_jaccard_summary.csv"))
cross_seed_N       <- read_csv(file.path(data_dir, "cross_seed_jaccard_vs_N.csv"))
gridpoints         <- read_csv(file.path(data_dir, "gridpoints.csv"))

# readr leaves the Python-style "True"/"False" as character — coerce to logical
equiv_sequential   <- equiv_sequential   |> mutate(converged = as.logical(converged))
stabilisation_seed <- stabilisation_seed |> mutate(all_points_stable = as.logical(all_points_stable))
cross_seed_summary <- cross_seed_summary |> mutate(is_n_max = as.logical(is_n_max))

N_MAX <- 500

# ---- Derived frames used by many chunks below (defined here so the chunks
#      work in any run order, not only top-to-bottom render) --------------
cross_at_max <- cross_seed_summary |> filter(is_n_max)

size_at_max <- equiv_sequential |>
  group_by(x, y) |>
  summarise(mean_equiv_size = mean(equiv_set_size_final),
            frac_seeds_converged = mean(converged), .groups = "drop")

map_df <- cross_at_max |> left_join(size_at_max, by = c("x", "y"))
```

## Background — what these two tests measure

For every **grid point** (a candidate landing spot on the fairway) and every
**sample size `N`** (Monte-Carlo approach shots simulated per club/aim option),
we compute the **equivalence set** `E*`: every (club, aim) combination whose
expected strokes-to-hole-out (ESHO) is within **one standard error** of the
single best option. `E*` is the set of strategies we *cannot statistically
separate* from optimal at that `N`.

- **Within-seed stabilisation (Test 1)** — does `E*` stop moving as we add
  shots? A grid point is declared *stabilised* at the first `N` where its `E*`is identical (Jaccard = 1) for **3 consecutive `N` steps**. Answers *how much simulation is enough*.
- **Cross-seed agreement (Test 2)** — 100 independent random seeds each run the
  full sweep. At a fixed `N`, how similar are the 100 `E*` sets they produce for
  the same grid point? Measured by **pairwise Jaccard**
  (`|intersection| / |union|`, averaged over all `100·99/2` seed pairs).
  Answers *whether the answer is reproducible* or a fluke of one seed.

Ideal end state: every point stabilised (Test 1) **and** every seed agreeing on
the same `E*` (Test 2 Jaccard → 1).

---

# Within-seed stabilisation

## % of points stabilised vs N EVER

```{r within-pct-ever, fig.width=9, fig.height=5.5}
ggplot(stabilisation_seed, aes(N, pct_stable_ever * 100, group = seed)) +
  geom_line(alpha = 0.15, colour = "steelblue") +
  stat_summary(aes(group = 1), fun = mean, geom = "line",
               linewidth = 1.3, colour = "black") +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "N (shots per grid point)",
       y = "% of grid points stabilised",
       title = "Within-seed stabilisation over N",
       subtitle = "thin lines = each of the 100 seeds, thick line = mean across seeds")
```

**What this shows.** For each seed, the running fraction of its 280 grid points that have met the "unchanged for 3 consecutive `N`" rule, vs `N`. 

**The statistic.** `pct_stable_ever` = (# points that have *ever* triggered the 3-in-a-row rule by this `N`) / (# points). Monotone non-decreasing by construction — a point that stabilises stays counted even if its `E*` later wobbles (the strict "still stable right now" version is `pct_stable_now` in `stabilisation_vs_N.csv`).

**Read it as.** This curve is the simulation-budget cost function: the `N` where the mean line flattens near 100% is the point of diminishing returns. A wide fan between the thin lines means the required budget is unpredictable run-to-run; a tight fan means it is well-determined.

## % of points stabilised vs N — NOW (the strict, non-cumulative version)

```{r within-pct-now, fig.width=9, fig.height=5.5}
ggplot(stabilisation_seed, aes(N, pct_stable_now * 100, group = seed)) +
  geom_line(alpha = 0.15, colour = "#D55E00") +
  stat_summary(aes(group = 1), fun = mean, geom = "line",
               linewidth = 1.3, colour = "black") +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "N (shots per grid point)",
       y = "% of grid points stable right now",
       title = "Within-seed stabilisation over N — NOW",
       subtitle = "same seeds as above, but pct_stable_now: is E* unchanged vs N-10, at this N specifically")
```

**Labelled for direct comparison with the EVER plot above.** `pct_stable_now`
is *not* cumulative and can go down as well as up — a point counts here only
if its `E*` is unchanged from the previous `N` step, right now. This is the
metric the "in plain terms" diagram further down is actually about: EVER
climbs to ~100%, NOW plateaus well below it, and the gap between the two
curves at large `N` *is* the flickering-tie phenomenon made visible as a
percentage rather than a single point's time series.

## Three stabilisation signals (mean across seeds)

```{r within-signals, fig.width=9, fig.height=5}
stabilisation_N |>
  select(N,
         `stabilised (E* unchanged 3x)` = mean_pct_stable_ever,
         `E* collapsed to a single option` = mean_pct_equiv_size1,
         `Jaccard of E* vs previous N` = mean_jaccard_vs_prev) |>
  pivot_longer(-N, names_to = "signal", values_to = "value") |>
  ggplot(aes(N, value * 100, colour = signal)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "N", y = "%", colour = NULL,
       title = "Within-seed convergence signals vs N (mean of 100 seeds)")
```

**What this shows.** Three ways of asking "has the model settled?", each averaged over the 100 seeds.

**The metrics.**

| line | metric | meaning |
|---|---|---|
| stabilised | `mean_pct_stable_ever` | % of points that passed the 3-consecutive-`N` Jaccard = 1 test |
| E\* size 1 | `mean_pct_equiv_size1` | % of points where `E*` currently holds exactly **one** club/aim option (a uniquely best play, no ties) |
| Jaccard vs prev | `mean_jaccard_vs_prev` | mean similarity of each point's `E*` between `N` and `N−10` (1 = no change this step) |

**Read it as.** "Jaccard vs prev" is the leading indicator — it rises first as
sets stop changing step-to-step. "Stabilised" lags it (needs 3 good steps in a
row). "E\* size 1" is a *different question*: a perfectly stable set can still
have size > 1 (a genuine tie). If "size 1" plateaus well below "stabilised",
that gap is real strategic ambiguity — more shots will not remove it.

## Three stabilisation signals — NOW (the strict, non-cumulative version)

```{r within-signals-now, fig.width=9, fig.height=5}
stabilisation_N |>
  select(N,
         `stable right now (E* == prev N)` = mean_pct_stable_now,
         `E* collapsed to a single option` = mean_pct_equiv_size1,
         `Jaccard of E* vs previous N` = mean_jaccard_vs_prev) |>
  pivot_longer(-N, names_to = "signal", values_to = "value") |>
  ggplot(aes(N, value * 100, colour = signal)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "N", y = "%", colour = NULL,
       title = "Within-seed convergence signals vs N — NOW version",
       subtitle = "'stabilised (ever)' swapped for 'stable right now' — everything else identical to the plot above")
```

**Labelled for direct comparison with the EVER version above.** Same three
signals, same seeds, same `N` axis — only the first line changed, from
"has this point *ever* passed the 3-consecutive test" to "is this point's
`E*` unchanged from the *immediately preceding* `N` step". Where the EVER
line keeps climbing toward 100%, the NOW line plateaus around 75% at
`N = 500` — noticeably *below* the Jaccard-vs-prev line (~93%), not close to
it: Jaccard is a continuous similarity score that gives partial credit for
near-matches, while "stable now" is a strict binary exact-match, so it's the
stricter and more honest of the two. That ~25-point gap between NOW and
100% at large `N` is the flickering described in the Conclusions and "In
plain terms" sections below, expressed as a population percentage rather
than a single point's time series.

## Per-seed: N at which the whole seed stabilised

```{r seed-conv-table}
seed_conv <- equiv_sequential |>
  group_by(seed) |>
  summarise(
    n_points       = n(),
    n_converged    = sum(converged),
    frac_converged = mean(converged),
    all_converged  = all(converged),
    all_stable_N   = if_else(all(converged), max(converged_N), NA_real_),
    .groups = "drop"
  )

seed_conv |>
  summarise(
    n_seeds                    = n(),
    seeds_fully_stabilised     = sum(all_converged),
    seeds_not_fully_stabilised = sum(!all_converged),
    `all_stable_N: median`     = median(all_stable_N, na.rm = TRUE),
    `all_stable_N: IQR`        = IQR(all_stable_N, na.rm = TRUE),
    `all_stable_N: min`        = min(all_stable_N, na.rm = TRUE),
    `all_stable_N: max`        = max(all_stable_N, na.rm = TRUE)
  ) |>
  pivot_longer(everything(), names_to = "statistic", values_to = "value") |>
  kable(digits = 1, caption = "Per-seed full-stabilisation summary")
```

**What this shows.** Each seed collapsed to one number — `all_stable_N`, the
largest `converged_N` over that seed's 280 points, i.e. the `N` at which the
*last straggler* point stabilised. Seeds with any never-stabilising point get
`NA` and land in `seeds_not_fully_stabilised`.

**The statistics.** `median / IQR / min / max` are over the seeds that *did*
fully stabilise. `IQR` (interquartile range) is the spread of the middle 50%.
`seeds_not_fully_stabilised` is the count that reached `N` = `r N_MAX` with ≥ 1
unstable point.

**Read it as.** `median all_stable_N` is the recommended simulation budget for a
typical run; `max` is the safe budget among seeds that converge at all. A
non-trivial `seeds_not_fully_stabilised` means the recommendation needs a
caveat — some runs always keep a few genuinely ambiguous points.

```{r seed-conv-hist, fig.width=8, fig.height=4.5}
ggplot(seed_conv, aes(all_stable_N)) +
  geom_histogram(binwidth = 10, boundary = 0, fill = "steelblue", colour = "white") +
  labs(x = "N at which every grid point in the seed had stabilised",
       y = "# seeds",
       title = "When does a whole seed stabilise?",
       subtitle = glue::glue("{sum(!seed_conv$all_converged)} of {nrow(seed_conv)} seeds never fully stabilised by N = {N_MAX}"))
```

**What this shows.** The distribution of `all_stable_N` across seeds
(never-fully-stabilised seeds are dropped — count in the subtitle).

**Read it as.** Tight and left-sited ⇒ the budget is small and reliable. A long
right tail ⇒ a few unlucky seeds drag the safe budget up. Mass piled against the
`N` = `r N_MAX` edge would mean the sweep stopped too early; clear empty space
before the edge means `N` = `r N_MAX` was long enough for essentially all
converging seeds.

## Equivalence-set size at stabilisation

```{r equiv-size, fig.width=8, fig.height=4.5}
converged_pts <- equiv_sequential |> filter(converged)

ggplot(converged_pts, aes(equiv_set_size_final)) +
  geom_bar(fill = "steelblue") +
  labs(x = "|E*| at the point of stabilisation",
       y = "# (grid point, seed) pairs",
       title = "How many club/aim options remain statistically tied once a point stabilises")

converged_pts |>
  count(equiv_set_size_final, name = "n_pairs") |>
  mutate(pct = round(100 * n_pairs / sum(n_pairs), 1)) |>
  kable(caption = "Distribution of |E*| at stabilisation")
```

**What this shows.** Among all (grid point, seed) pairs that stabilised, the
size of `E*` at the moment it stabilised — how many club/aim combinations are
still tied for best.

**The statistic.** `equiv_set_size_final` = `|E*|` at `converged_N`. The table
gives the count and % of pairs at each size.

**Read it as.** `|E*|` = 1 ⇒ one uniquely best strategy. `|E*|` ≥ 2 ⇒ the model
converged onto a genuine tie: several strategies are statistically
indistinguishable and would need a finer aim grid or a different metric to
separate (more Monte-Carlo shots will *not* help, because the set is already
stable). The mass above 1 is the irreducible strategic ambiguity of the hole.

---

# Cross-seed agreement

## Overall scorecard

```{r cross-overall}
cross_at_max <- cross_seed_summary |> filter(is_n_max)

tibble(
  metric = c(
    "grid points",
    "seeds",
    "% grid points converged (>=50% of seeds)      [overall summary]",
    "mean |E*| at convergence                      [overall summary]",
    "median |E*| at convergence                    [overall summary]",
    glue::glue("mean pairwise cross-seed Jaccard at N={N_MAX}"),
    glue::glue("median pairwise cross-seed Jaccard at N={N_MAX}"),
    glue::glue("grid points with FULL agreement at N={N_MAX} (mean J >= 0.999)"),
    glue::glue("grid points with mean J < 0.5 at N={N_MAX}")
  ),
  value = c(
    nrow(gridpoints),
    max(cross_seed_summary$n_seeds),
    equiv_overall$pct_grid_points_converged,
    round(equiv_overall$mean_equiv_set_size_at_convergence, 2),
    round(equiv_overall$median_equiv_set_size_at_convergence, 2),
    round(mean(cross_at_max$mean_jaccard), 3),
    round(median(cross_at_max$mean_jaccard), 3),
    sum(cross_at_max$mean_jaccard >= 0.999),
    sum(cross_at_max$mean_jaccard < 0.5)
  )
) |>
  kable(caption = "Headline numbers")
```

**What this shows.** A one-glance scorecard combining both tests at the end of
the sweep.

**The metrics.**

- **% grid points converged** — a point counts as converged if **≥ half** its
  seeds passed Test 1. `100` = every point converged for a majority of seeds.
- **mean / median `|E*|` at convergence** — over all converged (point, seed)
  pairs. `> 1` ⇒ the typical converged point still carries a strategic tie.
- **mean / median pairwise cross-seed Jaccard at `N` = `r N_MAX`** — averaged
  over seed pairs within a grid point, then over grid points. `1` ⇒ every seed
  produced an identical `E*` everywhere.
- **full agreement / mean J < 0.5** — the tails: how many points are completely
  reproducible vs essentially a coin-flip between seeds.

**Read it as.** Test 1 and Test 2 can disagree: a point can stabilise *within*
every seed yet different seeds stabilise onto *different* sets. High "%
converged" with a middling mean Jaccard is exactly that — locally stable,
globally not yet pinned down.

## Cross-seed Jaccard vs N

```{r cross-vs-n, fig.width=9, fig.height=5}
ggplot(cross_seed_N, aes(N, mean_jaccard_over_gridpoints)) +
  geom_line(linewidth = 1.2, colour = "firebrick") +
  geom_point(size = 1.6, colour = "firebrick") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "N (shots per grid point)",
       y = "mean pairwise Jaccard\n(averaged over all grid points)",
       title = "Do independent seeds converge to the same equivalence set?",
       subtitle = "1 = every seed produces an identical E* for every grid point")
```

**What this shows.** The headline Test 2 curve: at each `N`, the pairwise
cross-seed Jaccard averaged over all seed pairs and all grid points.

**The statistic.** `mean_jaccard_over_gridpoints` — for each grid point take the
mean of its `100·99/2` pairwise Jaccard values, then average those over the 280
grid points.

**Read it as.** Should climb monotonically and asymptote. The asymptote is the
*reproducibility ceiling*: if it levels off below 1, seeds keep disagreeing no
matter how long the sweep runs, because they are disagreeing about genuine ties,
not noise. Still rising at `N` = `r N_MAX` ⇒ more simulation would still buy
reproducibility.

## Cross-seed core vs union, and exact agreement

```{r cross-core-union, fig.width=9, fig.height=7}
p_core <- equiv_cross_seed |>
  group_by(N) |>
  summarise(`core (in EVERY seed)` = mean(core_size),
            `union (in ANY seed)`  = mean(union_size), .groups = "drop") |>
  pivot_longer(-N, names_to = "set", values_to = "size") |>
  ggplot(aes(N, size, colour = set)) +
  geom_line(linewidth = 1) +
  labs(x = "N", y = "mean set size over grid points", colour = NULL,
       title = "Core vs union of E* across seeds")

p_agree <- equiv_cross_seed |>
  group_by(N) |>
  summarise(mean_full_agreement = mean(full_agreement_frac), .groups = "drop") |>
  ggplot(aes(N, mean_full_agreement * 100)) +
  geom_line(linewidth = 1, colour = "firebrick") +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "N", y = "% of seeds equal to the core",
       title = "Fraction of seeds in exact agreement")

p_core / p_agree
```

**What this shows.** Two complementary views over `N`. *Top:* the **core** (`E*`
elements shared by *all* 100 seeds) and the **union** (elements in *any* seed),
each averaged over grid points. *Bottom:* the fraction of seeds whose `E*`
exactly equals the core.

**The metrics.** `core_size` = `|`intersection of all seed sets`|`.
`union_size` = `|`union of all seed sets`|`. `full_agreement_frac` = share of
seeds whose `E*` is identical to the core (adds nothing to the union).

**Read it as.** As seeds converge, core ↑ and union ↓ until they meet — the gap
between the two lines is the disagreement budget. `full_agreement_frac` → 1 is
the strongest reproducibility statement possible (not "similar" sets but *the
same* set). If core and union stay far apart, the average point has a solid
agreed core plus a fringe of seed-specific extras.

## Distribution of agreement across grid points (N = 500)

```{r cross-dist, fig.width=9, fig.height=4.5}
ggplot(cross_at_max, aes(mean_jaccard)) +
  geom_histogram(binwidth = 0.05, boundary = 0, fill = "firebrick", colour = "white") +
  labs(x = glue::glue("mean pairwise cross-seed Jaccard (N = {N_MAX})"),
       y = "# grid points",
       title = "Per-grid-point cross-seed agreement at the final N")

cross_at_max |>
  summarise(
    min = min(mean_jaccard), q25 = quantile(mean_jaccard, .25),
    median = median(mean_jaccard), mean = mean(mean_jaccard),
    q75 = quantile(mean_jaccard, .75), max = max(mean_jaccard)
  ) |>
  kable(digits = 3, caption = glue::glue("mean_jaccard across {nrow(cross_at_max)} grid points at N = {N_MAX}"))
```

**What this shows.** The spread behind the single "mean Jaccard" headline: one
value per grid point at `N` = `r N_MAX`, histogrammed, plus a five-number
summary.

**The statistics.** `min / q25 / median / mean / q75 / max` of the
per-grid-point `mean_jaccard`. `q25 / q75` are the 25th/75th percentiles (a
boxplot's box); a `mean` well below the `median` signals a left skew.

**Read it as.** A spike near 1 plus a spread of lower values (bimodal) is the
usual signature: many points fully reproducible, a distinct sub-population not.
The left tail names the problem points; the histogram mass below ~0.5 is the
share of the course where the optimal strategy *set* is not yet reproducible.

```{r cross-frac-identical, fig.width=9, fig.height=4.5}
ggplot(cross_at_max, aes(frac_pairs_identical)) +
  geom_histogram(binwidth = 0.05, boundary = 0, fill = "firebrick", colour = "white") +
  labs(x = glue::glue("fraction of seed-pairs with an identical E* (N = {N_MAX})"),
       y = "# grid points",
       title = "How often do two seeds land on exactly the same set?")
```

**What this shows.** For each grid point, the share of the `100·99/2` seed pairs
whose `E*` sets are *exactly* equal (Jaccard exactly 1), at `N` = `r N_MAX`.

**The statistic.** `frac_pairs_identical` — a stricter cousin of `mean_jaccard`:
only perfect matches count, "close" doesn't.

**Read it as.** This is the number to quote for "reproducible". A point with
`mean_jaccard` = 0.9 but `frac_pairs_identical` = 0.1 has sets that are
*similar* but rarely *identical* — usually one flip-flopping fringe element.
Points near `frac_pairs_identical` = 1 are locked in.

## Best- and worst-agreeing grid points (N = 500)

```{r cross-extremes}
cross_at_max |>
  arrange(desc(mean_jaccard)) |>
  slice_head(n = 10) |>
  select(gp_id, x, y, mean_jaccard, median_jaccard, min_jaccard, frac_pairs_identical) |>
  kable(digits = 3, caption = "Top 10 grid points — seeds agree most")

cross_at_max |>
  arrange(mean_jaccard) |>
  slice_head(n = 10) |>
  select(gp_id, x, y, mean_jaccard, median_jaccard, min_jaccard, frac_pairs_identical) |>
  kable(digits = 3, caption = "Bottom 10 grid points — seeds disagree most")
```

**What this shows.** The named extremes: the 10 most and 10 least reproducible
grid points, with hole coordinates.

**The columns.** `mean_jaccard / median_jaccard` — average and typical pairwise
similarity; `min_jaccard` — the worst single seed pair (0 = two seeds share no
strategy at all); `frac_pairs_identical` — share of exact matches.

**Read it as.** Cross-reference the bottom-10 `(x, y)` against the spatial maps
below. If the worst points cluster in one region of the hole (a wide fairway
landing zone, a hazard edge) that is a substantive finding about the course, not
a numerical artefact.

## Side-by-side hole maps: cross-seed agreement vs equivalence-set size (N = 500)

```{r cross-spatial-side, fig.width=13, fig.height=8}
size_at_max <- equiv_sequential |>
  group_by(x, y) |>
  summarise(mean_equiv_size = mean(equiv_set_size_final),
            frac_seeds_converged = mean(converged), .groups = "drop")

map_df <- cross_at_max |> left_join(size_at_max, by = c("x", "y"))

p_j <- ggplot(map_df, aes(x, y, fill = mean_jaccard)) +
  geom_tile(width = 11, height = 11) +
  scale_fill_viridis_c(limits = c(0, 1), name = "mean\nJaccard") +
  coord_equal() +
  labs(x = "x (yards)", y = "y (yards)",
       title = "Cross-seed agreement",
       subtitle = "yellow = seeds agree, purple = they don't")

p_s <- ggplot(map_df, aes(x, y, fill = mean_equiv_size)) +
  geom_tile(width = 11, height = 11) +
  scale_fill_viridis_c(option = "magma", direction = -1, name = "mean\n|E*|") +
  coord_equal() +
  labs(x = "x (yards)", y = "y (yards)",
       title = "Equivalence-set size",
       subtitle = "dark = one clear best play, light = many tied options")

p_j + p_s
```

**What this shows.** The same 280 grid points on the hole, coloured two ways at
`N` = `r N_MAX`. **Left:** mean pairwise cross-seed Jaccard — *do the seeds agree
here?* **Right:** mean `|E*|` across seeds — *how many strategies are tied here?*

**The statistics.** Left fill = per-point `mean_jaccard` (0–1). Right fill =
per-point mean of `equiv_set_size_final` over the 100 seeds. Tiles ≈ 11 yd, the
grid spacing.

**Read it as.** The two maps should be near mirror images: **large `|E*|` (light,
right) lines up with low agreement (purple, left)**. That is the mechanism —
where many club/aim options are genuinely tied, each seed's noise picks a
slightly different subset, so pairwise Jaccard drops. Dark-right / yellow-left
regions are solved: one best play, every seed finds it. Disagreement (purple
left) where the set is *small* (dark right) is anomalous — likely a knife-edge
between two distinct strategies rather than a broad tie — and worth a look.

## Does within-seed ambiguity explain cross-seed disagreement?

```{r cross-vs-size, fig.width=8, fig.height=5}
ggplot(map_df, aes(mean_equiv_size, mean_jaccard)) +
  geom_point(aes(colour = frac_seeds_converged), size = 2, alpha = 0.8) +
  scale_colour_viridis_c(name = "frac seeds\nstabilised") +
  labs(x = "mean |E*| across seeds at N = 500",
       y = "mean pairwise cross-seed Jaccard",
       title = "Larger tied-best sets -> seeds agree less")
```

**What this shows.** The relationship in the two maps, as a scatter: one point
per grid point, `|E*|` size on x, cross-seed Jaccard on y, coloured by the
fraction of seeds that stabilised there.

**Read it as.** A clear downward trend confirms the mechanism — cross-seed
disagreement is driven by genuine strategic ambiguity (`|E*|` > 1), not by
under-sampling. Bottom-right points (big set, low agreement, often low
`frac_seeds_converged`) are the hard cases; top-left (size ≈ 1, Jaccard ≈ 1) are
fully solved. Vertical scatter at a fixed small `|E*|` would mean something other
than set size is causing disagreement there.

## Grid points that never fully stabilised

```{r nonconv-table}
equiv_nonconv |>
  arrange(desc(frac_seeds_not_converged)) |>
  kable(digits = 3,
        caption = "Grid points with >=1 seed that hit N = 500 without a stable E*")
```

**What this shows.** Every grid point where at least one seed reached
`N` = `r N_MAX` with an `E*` still changing.

**The columns.** `n_seeds_not_converged / frac_seeds_not_converged` — how many /
what share of the 100 seeds failed to stabilise here.
`mean_equiv_set_size_stuck / max_equiv_set_size_stuck` — the size of the
still-moving `E*` for those seeds (large = a wide, genuinely ambiguous landing
zone; small = a persistent flip between two near-identical options).

**Read it as.** The definitive list of "hard" grid points. If
`frac_seeds_not_converged` is a few percent everywhere, the sweep is essentially
complete and these are edge cases. Large stuck sizes mark regions where the
recommendation should be reported as a *set* of acceptable plays, not a single
optimum.

---

# Anatomy of ambiguous equivalence sets

```{r ambig-load}
eq_files <- list.files(outputs_dir, "seed[0-9]+_N0500_equivset\\.csv$",
                       recursive = TRUE, full.names = TRUE)

eq500 <- map_dfr(eq_files, read_csv, show_col_types = FALSE) |>
  mutate(is_argmin = as.logical(is_argmin))

# per (grid point, seed): how spread out is this seed's E* ?
set_spread <- eq500 |>
  group_by(x, y, seed) |>
  summarise(
    set_size    = n(),
    n_clubs     = n_distinct(club),
    esho_range  = max(esho_mean) - min(esho_mean),          # strokes, best..worst in E*
    aim_span    = max(aim_offset) - min(aim_offset),        # yards
    SE_min      = first(SE_min),
    esho_range_rel = (max(esho_mean) - min(esho_mean)) / first(SE_min),  # 0..1 of the 1-SE band used
    .groups = "drop"
  )

# aggregate to the grid point and tag by cross-seed agreement
point_spread <- set_spread |>
  group_by(x, y) |>
  summarise(across(c(set_size, n_clubs, esho_range, aim_span, esho_range_rel), mean),
            .groups = "drop") |>
  left_join(select(cross_at_max, x, y, mean_jaccard, frac_pairs_identical),
            by = c("x", "y")) |>
  mutate(agreement = factor(
    case_when(
      mean_jaccard >= 0.80 ~ "clear (J >= 0.80)",
      mean_jaccard >= 0.50 ~ "mixed (0.50-0.80)",
      TRUE                 ~ "ambiguous (J < 0.50)"
    ),
    levels = c("clear (J >= 0.80)", "mixed (0.50-0.80)", "ambiguous (J < 0.50)")
  ))

point_spread |> count(agreement) |> kable(caption = "Grid points by cross-seed agreement class")
```

## How wide is the equivalence set — ambiguous vs clear points

```{r ambig-boxes, fig.width=11, fig.height=8}
metric_labels <- c(
  set_size       = "|E*|  (# tied club/aim options)",
  n_clubs        = "# distinct clubs in E*",
  aim_span       = "aim span in E*  (yards, max - min)",
  esho_range_rel = "ESHO range in E*  (fraction of the 1-SE band)"
)

point_spread |>
  select(agreement, all_of(names(metric_labels))) |>
  pivot_longer(-agreement, names_to = "metric", values_to = "value") |>
  mutate(metric = factor(metric_labels[metric], levels = metric_labels)) |>
  ggplot(aes(agreement, value, fill = agreement)) +
  geom_violin(alpha = 0.4, colour = NA) +
  geom_boxplot(width = 0.18, outlier.size = 0.6) +
  facet_wrap(~ metric, scales = "free_y") +
  scale_fill_manual(values = c("#2e7d32", "#f9a825", "#c62828"), guide = "none") +
  labs(x = NULL, y = NULL,
       title = "What the equivalence set looks like, by cross-seed agreement class",
       subtitle = "each point = one grid point, averaged over its 100 seeds")
```

**What this shows.** Four descriptors of the equivalence set, split by how well
the seeds agree at that grid point (green / amber / red = clear / mixed /
ambiguous, thresholds on `mean_jaccard`). Each dot is a grid point; the value is
the mean over its 100 seeds.

**The metrics.**

- **`|E*|`** — number of (club, aim) options that survive the 1-SE cut.
- **`# distinct clubs`** — how many *different clubs* are in the set (2 aims of
  the same club counts as 1). Distinguishes "same club, fine aim wobble" from
  "genuinely different clubs are tied".
- **`aim span`** — `max(aim_offset) − min(aim_offset)` across the set, in yards.
- **`ESHO range (fraction of 1-SE band)`** — `(max ESHO − min ESHO) / SE_min`.
  By construction this is between 0 and 1; near 1 means the set stretches the
  full standard error, near 0 means every member has almost the same expected
  score.

**Read it as.** If the red group sits higher on `|E*|`, `# clubs` and `aim span`
but its **ESHO range is still small** (all members score within a hair of each
other), the disagreement is *benign*: the seeds are quibbling between plays that
are, for scoring purposes, interchangeable. If the red group also has a large
ESHO range, the seeds are disagreeing about options that are measurably
different — a more serious reproducibility problem.

## ESHO spread inside the set, vs cross-seed agreement

```{r ambig-scatter, fig.width=9, fig.height=5.5}
ggplot(point_spread, aes(mean_jaccard, esho_range, colour = n_clubs)) +
  geom_point(size = 2, alpha = 0.85) +
  scale_colour_viridis_c(name = "# clubs\nin E*") +
  labs(x = "mean pairwise cross-seed Jaccard (N = 500)",
       y = "ESHO range within E*  (strokes, best - worst member)",
       title = "Do low-agreement points also have a wide range of expected scores?",
       subtitle = "y is capped at SE_min by construction")
```

**What this shows.** One point per grid point: cross-seed agreement on the x
axis, the spread of expected strokes *inside* the equivalence set on the y axis,
coloured by how many different clubs are tied.

**Read it as.** A negative slope means low-agreement points genuinely span a
wider band of expected outcomes. If instead the cloud is flat and low
(everything within ~0.02 strokes regardless of Jaccard) then even the worst
disagreement is between near-identical plays — quote that number as the "cost of
irreproducibility in strokes". Colour tells you whether wide sets are one club
with many aims (low # clubs) or a true multi-club tie (high # clubs).

## What is actually tied — the 9 least reproducible grid points

```{r ambig-composition, fig.width=12, fig.height=9}
worst9 <- cross_at_max |> arrange(mean_jaccard) |> slice_head(n = 9) |>
  mutate(panel = glue::glue("gp {gp_id}  ({round(x)}, {round(y)})  J={round(mean_jaccard,2)}"))

eq500 |>
  inner_join(select(worst9, x, y, panel), by = c("x", "y")) |>
  count(panel, club, aim_offset, name = "n_seeds") |>
  group_by(panel) |>
  mutate(club = fct_reorder(club, n_seeds, sum)) |>
  ungroup() |>
  ggplot(aes(aim_offset, club, size = n_seeds, colour = n_seeds)) +
  geom_point(alpha = 0.8) +
  facet_wrap(~ panel, scales = "free_y") +
  scale_size_area(max_size = 7, name = "# seeds") +
  scale_colour_viridis_c(name = "# seeds") +
  labs(x = "aim offset (yards)", y = NULL,
       title = "Equivalence-set membership across seeds, at the 9 worst grid points",
       subtitle = "a dot = a (club, aim) that appeared in E* for that many of the 100 seeds")
```

**What this shows.** For each of the 9 lowest-Jaccard grid points, every
(club, aim) combination that showed up in *any* seed's `E*`, with dot size/colour
= how many of the 100 seeds included it.

**Read it as.** A few big dots + many small ones = there is a stable agreed core
(the big dots, in nearly every seed) plus a scatter of seed-specific fringe
picks — the fringe is what drives Jaccard down. Big dots spread across *several
clubs* (rows) means genuinely different clubs are competitive here; big dots in
*one row across many aim values* means it is one club and the aim is
under-determined. Vertically compact clusters are easier to summarise as a
recommendation than spread-out ones.

## Is the ambiguity driven by club or by aim?

For each grid point we look at every seed's **single best (arg-min) combination**
and ask: do the seeds disagree about which *club* to hit, or do they agree on
club and only disagree about the *aim offset*?

```{r ambig-driver-setup}
argmins <- eq500 |>
  filter(is_argmin) |>
  distinct(x, y, seed, .keep_all = TRUE) |>
  select(x, y, seed, argmin_club = club, argmin_aim = aim_offset)

driver_df <- argmins |>
  group_by(x, y) |>
  mutate(modal_club = { t <- table(argmin_club); names(t)[which.max(t)] }) |>
  summarise(
    n_seeds          = n(),
    club_instability = mean(argmin_club != modal_club),          # frac of seeds off the most-common best club
    aim_sd_modal     = sd(argmin_aim[argmin_club == modal_club]),# sd of best aim (yд), among seeds that agree on club
    n_best_clubs     = n_distinct(argmin_club),
    .groups = "drop"
  ) |>
  left_join(select(cross_at_max, x, y, mean_jaccard), by = c("x", "y")) |>
  mutate(
    aim_sd_modal = replace_na(aim_sd_modal, 0),
    driver = factor(
      case_when(
        mean_jaccard >= 0.80                             ~ "reproducible",
        club_instability >= 0.15 & aim_sd_modal >= 2.5   ~ "club & aim",
        club_instability >= 0.15                         ~ "club",
        aim_sd_modal >= 1.0                              ~ "aim",
        TRUE                                             ~ "minimal"
      ),
      levels = c("reproducible", "minimal", "aim", "club", "club & aim")
    )
  )
```

```{r ambig-driver-scatter, fig.width=9, fig.height=6}
lab <- tibble(
  club_instability = c(0.02, 0.35, 0.02, 0.35),
  aim_sd_modal     = c(0.6, 0.6, 11, 11),
  text = c("reproducible-ish", "club-driven", "aim-driven", "club & aim")
)

driver_df |>
  filter(mean_jaccard < 0.80) |>
  ggplot(aes(club_instability, aim_sd_modal)) +
  geom_vline(xintercept = 0.15, linetype = 2, colour = "grey55") +
  geom_hline(yintercept = 2.5, linetype = 2, colour = "grey55") +
  geom_point(aes(colour = mean_jaccard, size = n_best_clubs), alpha = 0.85) +
  geom_text(data = lab, aes(label = text), colour = "grey35", fontface = "italic", size = 3.5) +
  scale_colour_viridis_c(limits = c(0, 1), name = "cross-seed\nJaccard") +
  scale_size_continuous(name = "# distinct\nbest clubs", range = c(1, 5)) +
  labs(x = "club instability  (fraction of seeds NOT on the modal best club)",
       y = "aim instability  (SD of best aim across seeds, yards)",
       title = "What makes the seeds disagree at each ambiguous grid point?",
       subtitle = "non-reproducible points only (mean Jaccard < 0.80)")
```

**What this shows.** One dot per non-reproducible grid point. **x** = how often
the seeds pick a *different club* as best (0 = every seed agrees on the club).
**y** = how much the *best aim* wanders between seeds (standard deviation in
yards, computed only over the seeds that agree on the club, so it is a pure aim
signal). Colour = cross-seed Jaccard; size = number of distinct clubs that were
ever "best" here.

**The metrics.** `club_instability` = `mean(argmin_club != modal_club)` across
the 100 seeds. `aim_sd_modal` = `sd(argmin_aim)` restricted to seeds on the
modal club. Dashed lines at `0.15` and `2.5 yд` (half the 5 yд aim grid) split
the plane into the four regimes labelled in grey.

**Read it as.** Bottom-right = *club-driven*: seeds can't agree which club to
hit. Top-left = *aim-driven*: they agree on the club, the aim is
under-determined. Top-right = both. Bottom-left = barely ambiguous at all. Where
the cloud sits tells you what a finer grid would need to fix — more aim
resolution vs a tie-break rule between clubs.

```{r ambig-driver-bar, fig.width=8, fig.height=4}
driver_df |>
  count(driver) |>
  mutate(pct = round(100 * n / sum(n), 1)) |>
  ggplot(aes(n, fct_rev(driver), fill = driver)) +
  geom_col() +
  geom_text(aes(label = glue::glue("{n}  ({pct}%)")), hjust = -0.1, size = 3.4) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  scale_fill_manual(values = c(reproducible = "#2e7d32", minimal = "#9e9e9e",
                               aim = "#1565c0", club = "#c62828",
                               `club & aim` = "#6a1b9a"), guide = "none") +
  labs(x = "# grid points", y = NULL,
       title = "Every grid point classified by what drives its ambiguity")

driver_df |>
  count(driver, name = "n_grid_points") |>
  mutate(pct = round(100 * n_grid_points / sum(n_grid_points), 1)) |>
  kable(caption = "Ambiguity driver — grid-point counts")
```

**What this shows.** All 280 grid points bucketed by the rule in the setup
chunk: `reproducible` (Jaccard ≥ 0.80), then `club` / `aim` / `club & aim`
depending on which instability crosses its threshold, else `minimal`.

**Read it as.** This is the one-line answer to "is it club or aim?". If the
`aim` bar dominates, the equivalence-set ambiguity is almost entirely about aim
resolution — the model knows the club, and a finer aim grid (or reporting aim as
a tolerance band) would resolve most of it. A large `club` bar would instead
mean genuinely different clubs are tied and the recommendation has to name a
set. `club & aim` points are the messiest and worth listing individually.

---

# Course-context maps

The same results with the hole drawn underneath, so location on the hole is
legible. Geometry comes from `export_hole_geometry.py` (same game coordinates as
the strategy grid).

```{r course-setup}
geom_raw   <- read_csv(file.path(data_dir, "hole_geometry.csv"), show_col_types = FALSE)
hole_sf    <- st_as_sf(geom_raw, wkt = "wkt")
hole_polys <- hole_sf |> filter(!lie %in% c("tee", "pin"))
hole_marks <- hole_sf |> filter(lie %in% c("tee", "pin"))

ob_bounds  <- read_csv(file.path(data_dir, "ob_bounds.csv"), show_col_types = FALSE) |>
  deframe()   # named vector: ob_x_left, ob_x_right, ob_y_far

# Fixed view spanning all three OB lines — grid points only reach y ~284, well
# short of ob_y_far, so without this the "past the bunkers" line would be
# clipped out of view on any plot built from grid-point data.
hole_bbox <- st_bbox(hole_sf)
plot_xlim <- c(ob_bounds["ob_x_left"] - 5, ob_bounds["ob_x_right"] + 5)
plot_ylim <- c(min(hole_bbox["ymin"], 0) - 5, ob_bounds["ob_y_far"] + 10)

lie_fill <- c(green = "#a8d5a2", fairway = "#8bc34a",
              rough = "#5d8a3a", water = "#5db7e8", bunker = "#e8dcae")

hole_layer <- function() {
  list(
    # OB regions first, so course geometry draws on top
    annotate("rect", xmin = -Inf, xmax = ob_bounds["ob_x_left"], ymin = -Inf, ymax = Inf,
             fill = "firebrick", alpha = 0.15),
    annotate("rect", xmin = ob_bounds["ob_x_right"], xmax = Inf, ymin = -Inf, ymax = Inf,
             fill = "firebrick", alpha = 0.15),
    annotate("rect", xmin = -Inf, xmax = Inf, ymin = ob_bounds["ob_y_far"], ymax = Inf,
             fill = "firebrick", alpha = 0.15),
    geom_vline(xintercept = c(ob_bounds["ob_x_left"], ob_bounds["ob_x_right"]),
               colour = "firebrick", linetype = 2, linewidth = 0.4),
    geom_hline(yintercept = ob_bounds["ob_y_far"], colour = "firebrick", linetype = 2, linewidth = 0.4),
    geom_sf(data = hole_polys, aes(fill = lie), colour = NA, alpha = 0.85,
            inherit.aes = FALSE),
    geom_sf(data = hole_marks, aes(shape = lie), size = 2.6, stroke = 1,
            colour = "black", fill = "white", inherit.aes = FALSE),
    scale_fill_manual(values = lie_fill, name = "lie", guide = guide_legend(order = 2)),
    scale_shape_manual(values = c(tee = 15, pin = 21), name = NULL),
    coord_sf(xlim = plot_xlim, ylim = plot_ylim, expand = FALSE),
    theme(axis.title = element_blank())
  )
}
```

## Cross-seed agreement and equivalence-set size, on the hole

```{r course-maps, fig.width=13, fig.height=9}
pj <- ggplot() + hole_layer() +
  geom_point(data = map_df, aes(x, y, colour = mean_jaccard), size = 2.6) +
  scale_colour_viridis_c(limits = c(0, 1), name = "cross-seed\nJaccard",
                         guide = guide_colourbar(order = 1)) +
  labs(title = "Where do the seeds agree?")

ps <- ggplot() + hole_layer() +
  geom_point(data = map_df, aes(x, y, colour = mean_equiv_size), size = 2.6) +
  scale_colour_viridis_c(option = "magma", direction = -1, name = "mean |E*|",
                         guide = guide_colourbar(order = 1)) +
  labs(title = "Where are the tied-best sets large?")

pj + ps
```

**What this shows.** The two headline per-grid-point quantities from the "side by
side hole maps" section, now overlaid on the actual hole — green, fairway,
rough, water, bunkers, tee (square) and pin (circle).

**Read it as.** Now you can attribute the pattern to course features: e.g.
low-agreement / large-`|E*|` points sitting in the wide part of the fairway
(many landing spots score the same), or hugging a water margin (a knife-edge
where a few yards of aim flip the outcome). Points over or just short of the
green will read differently from points 150+ yд out.

## The ambiguous cluster and the never-stabilised points

```{r course-hotspots, fig.width=9, fig.height=10}
nonconv_pts <- equiv_nonconv |> distinct(x, y) |> mutate(kind = "never stabilised")

ggplot() + hole_layer() +
  geom_point(data = point_spread, aes(x, y, colour = agreement), size = 2.8) +
  geom_point(data = nonconv_pts, aes(x, y), shape = 4, size = 3.5, stroke = 1.1,
             colour = "black") +
  scale_colour_manual(values = c("clear (J >= 0.80)" = "#2e7d32",
                                 "mixed (0.50-0.80)" = "#f9a825",
                                 "ambiguous (J < 0.50)" = "#c62828"),
                      name = "agreement class") +
  labs(title = "Ambiguity classes on the hole",
       subtitle = "x = grid point where >=1 seed never stabilised by N = 500")
```

**What this shows.** Every grid point coloured by agreement class (green / amber
/ red), with a black **×** on the points that also failed Test 1 for at least one
seed.

**Read it as.** If the red points and the ×'s form a compact cluster in one area,
that region is the single driver of both non-reproducibility and
non-stabilisation — worth describing explicitly in the write-up (what is there:
hazard, fairway width, distance band). Scattered red points with no × are
"stably ambiguous": each seed settles, they just settle differently.

## Equivalence-set aim spread and landing spots

**Data provenance, to be explicit:** every plot in this subsection is built
from `eq500` — the actual HPC output, `outputs/seed*/seed*_N0500_equivset.csv`
for all **100 seeds**, pooled together at `N = 500`. It is not one seed's run;
`n_seeds` in each plot literally counts how many of the 100 independent seeds
included that exact (club, aim) in their equivalence set. Nothing here is
simulated fresh — it's a re-read of the real convergence-study output.

### Aim-offset fan (by angle) — the 9 worst grid points

```{r course-aimspread-rays, fig.width=9, fig.height=10}
worst_members_rays <- eq500 |>
  inner_join(select(worst9, x, y), by = c("x", "y")) |>
  group_by(x, y, club, aim_offset) |>
  summarise(n_seeds = n(), .groups = "drop") |>
  filter(n_seeds >= 5)

ggplot() + hole_layer() +
  geom_segment(data = worst_members_rays,
               aes(x = x, y = y, xend = x + aim_offset, yend = y + 12,
                   colour = club, alpha = n_seeds),
               linewidth = 0.7) +
  geom_point(data = distinct(worst_members_rays, x, y),
             aes(x, y), colour = "black", size = 2) +
  scale_alpha(range = c(0.25, 1), name = "# seeds") +
  labs(title = "Aim spread of the equivalence set at the 9 worst grid points",
       subtitle = "each ray = a (club, aim) tied for best; length along x is the aim offset (yards)")
```

**What this shows.** For the 9 least-reproducible grid points, a fan of short
rays from the point — one per (club, aim) that appears in ≥ 5 seeds' `E*` —
coloured by club, faded by how many seeds picked it. The horizontal reach of a
ray is its aim offset in yards, not a real landing position.

**Read it as.** A tight fan of one colour = the seeds basically agree on club and
are only nudging aim; a wide multi-colour fan = genuinely different plays are all
live.

⚠️ **This schematic is directionally misleading and shouldn't be read as "which
way the shot goes."** `xend = x + aim_offset` is a naive offset — it does not
account for the true bearing from the grid point to the pin. Checked directly:
for **46 of these 132 (club, aim) combos (35%)**, the naive ray points
left/right **opposite** to where the shot actually lands once the real
rotation toward the pin is applied (e.g. the pin sits at `x=5`, so a point at
`x=-40` needs a large rightward rotation just to aim at the pin at all — a
positive `aim_offset` there can still finish left of where this schematic
draws it). The two "worst 9" plots you compared **are the same 132 rows of
data** (verified identical) — only this ray's endpoint is wrong. The corrected
version below fixes that.

```{r course-aimspread-setup}
worst_members <- eq500 |>
  inner_join(select(worst9, x, y), by = c("x", "y")) |>
  group_by(x, y, club, aim_offset) |>
  summarise(n_seeds = n(), .groups = "drop") |>
  filter(n_seeds >= 5)   # drop 1-off fringe picks for legibility

club_means <- read_csv("../../data/simulated_lpga_shot_data2.csv", show_col_types = FALSE) |>
  group_by(Club) |>
  summarise(mean_side = mean(Side), mean_carry = mean(Carry), .groups = "drop")

pin_xy <- hole_marks |> filter(lie == "pin") |> st_coordinates()
pin_x  <- pin_xy[1, "X"]; pin_y <- pin_xy[1, "Y"]

# same rotation as core.py::rotation_translator, applied to the club's MEAN
# (Side, Carry) instead of a Monte Carlo draw -> the average landing spot
worst_landing <- worst_members |>
  left_join(club_means, by = c("club" = "Club")) |>
  mutate(
    total_dist = sqrt((pin_x - x)^2 + (pin_y - y)^2),
    angle_rad  = if_else(total_dist > 0, atan(aim_offset / total_dist), 0),
    rot_x  = cos(angle_rad) * mean_side - sin(angle_rad) * mean_carry,
    rot_y  = sin(angle_rad) * mean_side + cos(angle_rad) * mean_carry,
    unit_x = (pin_x - x) / total_dist,
    unit_y = (pin_y - y) / total_dist,
    land_x = x + unit_x * rot_y - unit_y * rot_x,
    land_y = y + unit_y * rot_y + unit_x * rot_x,
  )
```

### Aim-offset fan, corrected (true bearing) — the 9 worst grid points

```{r course-aimspread-rays-fixed, fig.width=9, fig.height=10}
worst_direction <- worst_landing |>
  mutate(
    vec_x = land_x - x, vec_y = land_y - y,
    vec_len = sqrt(vec_x^2 + vec_y^2),
    short_x = x + vec_x / vec_len * 15,   # truncated to a fixed 15 yd stub,
    short_y = y + vec_y / vec_len * 15,   # same visual scale as the naive fan above
  )

ggplot() + hole_layer() +
  geom_segment(data = worst_direction,
               aes(x = x, y = y, xend = short_x, yend = short_y, colour = club, alpha = n_seeds),
               linewidth = 0.7) +
  geom_point(data = distinct(worst_direction, x, y),
             aes(x, y), colour = "black", size = 2) +
  scale_alpha(range = c(0.25, 1), name = "# seeds") +
  labs(title = "Aim spread of the equivalence set, true bearing toward the pin",
       subtitle = "same 132 (club, aim) rows as the schematic above — this time the ray direction is real")
```

**What this shows.** The identical data as the naive fan, but each ray is now
a 15 yd stub cut from the *actual* rotated landing vector (the same
computation the landing-spot plot uses), not `x + aim_offset`. Same colours,
same rows — the direction is now trustworthy, and this ray is exactly the
first 15 yards of the corresponding dot's segment in the plot below.

**Read it as.** Compare this to the naive version above at the same 9 points —
notice several fans visibly flip orientation once the true bearing is applied
(most obviously at points far from `x = 5`, where the naive schematic's
left/right has the least in common with reality). Trust this one, and the
"Mean landing spot" plot next — not the first ray-fan.

### Mean landing spot — the 9 worst grid points

Rather than just the aim *angle*, this computes the real mean landing spot of
each tied-best (club, aim) — using that club's average (Side, Carry) from the
LPGA shot data (`simulated_lpga_shot_data2.csv`), rotated the same way
`core.py::rotation_translator` does — so you can see where the "average"
outcome of each competing option actually finishes on the course. (`worst_landing`
was already computed above, for the corrected fan — reused here as-is.)

```{r course-aimspread, fig.width=9, fig.height=10}
ggplot() + hole_layer() +
  geom_segment(data = worst_landing,
               aes(x = x, y = y, xend = land_x, yend = land_y, colour = club, alpha = n_seeds),
               linewidth = 0.4) +
  geom_point(data = worst_landing,
             aes(land_x, land_y, colour = club, size = n_seeds), alpha = 0.85) +
  geom_point(data = distinct(worst_landing, x, y),
             aes(x, y), colour = "black", size = 2, shape = 17) +
  scale_alpha(range = c(0.2, 0.7), guide = "none") +
  scale_size_continuous(range = c(1.5, 4), name = "# seeds") +
  labs(title = "Mean landing spot of every tied-best (club, aim), at the 9 worst grid points",
       subtitle = "triangle = the grid point itself; dot = where that club/aim lands on average; thin line traces dot back to its origin")
```

**What this shows.** For the 9 least-reproducible grid points (black triangles),
every (club, aim) that appears in ≥ 5 seeds' `E*`, plotted not as an aim angle
but as its actual **mean landing position** — the club's average carry and side
miss (from the raw LPGA data), rotated toward the pin by that aim offset,
exactly like the real shot simulation does, just using the mean instead of a
random draw. Colour = club, dot size = how many seeds picked it, thin line =
which origin point it belongs to (several origins are close together on this
hole, so the traceback line matters).

**Read it as.** This is the direct answer to "if I just play any member of
`E*`, where does the ball end up?" A tight cluster of dots = the tied options
all land in basically the same spot regardless of which one is picked — safe
to treat as interchangeable. Dots spread across the fairway/rough boundary, or
near the water/OB shading, mean the "tied" options are not equally safe *in
practice* even though they score the same in expectation — worth a second look
before treating `E*` as freely interchangeable advice.

### Mean landing spot — every grid point, batched by row

All 280 grid points at once would be an unreadable smear, so this walks the
strategy grid **3 y-rows at a time** (28 rows -> 10 batches, ~30 origins and
~100-250 dots each), same landing-spot construction as above, applied to every
point rather than just the worst 9. Same `n_seeds >= 5` legibility filter — a
grid point with no dots in its batch means every seed's pick there got fewer
than 5 votes each (i.e. total disagreement across many near-unique choices,
itself worth noting).

```{r course-aimspread-all-batches, fig.width=9, fig.height=10}
all_pts <- distinct(eq500, x, y)
y_vals  <- sort(unique(all_pts$y))
row_batches <- split(y_vals, ceiling(seq_along(y_vals) / 3))

for (b in seq_along(row_batches)) {
  batch_y <- row_batches[[b]]

  batch_members <- eq500 |>
    filter(y %in% batch_y) |>
    group_by(x, y, club, aim_offset) |>
    summarise(n_seeds = n(), .groups = "drop") |>
    filter(n_seeds >= 5)

  batch_landing <- batch_members |>
    left_join(club_means, by = c("club" = "Club")) |>
    mutate(
      total_dist = sqrt((pin_x - x)^2 + (pin_y - y)^2),
      angle_rad  = if_else(total_dist > 0, atan(aim_offset / total_dist), 0),
      rot_x  = cos(angle_rad) * mean_side - sin(angle_rad) * mean_carry,
      rot_y  = sin(angle_rad) * mean_side + cos(angle_rad) * mean_carry,
      unit_x = (pin_x - x) / total_dist,
      unit_y = (pin_y - y) / total_dist,
      land_x = x + unit_x * rot_y - unit_y * rot_x,
      land_y = y + unit_y * rot_y + unit_x * rot_x,
    )

  p <- ggplot() + hole_layer() +
    geom_segment(data = batch_landing,
                 aes(x = x, y = y, xend = land_x, yend = land_y, colour = club, alpha = n_seeds),
                 linewidth = 0.3) +
    geom_point(data = batch_landing,
               aes(land_x, land_y, colour = club, size = n_seeds), alpha = 0.8) +
    geom_point(data = distinct(batch_landing, x, y),
               aes(x, y), colour = "black", size = 1.6, shape = 17) +
    scale_alpha(range = c(0.15, 0.6), guide = "none") +
    scale_size_continuous(range = c(1, 3.5), name = "# seeds") +
    labs(title = glue::glue("Mean landing spots — rows y = {round(min(batch_y))}-{round(max(batch_y))}  (batch {b}/{length(row_batches)})"),
         subtitle = "triangle = grid point; dot = mean landing spot of a tied-best (club, aim), colour = club")

  print(p)
}
```

**What this shows.** The same mean-landing-spot construction as the worst-9
plot, run for every one of the 280 grid points, split into 10 panels by
strategy-grid row so each panel stays legible. Scanning top to bottom moves
from the tee-ward rows to the green-ward rows.

**Read it as.** Compare panels: rows where the equivalence-set dots cluster
tightly (little colour variety, landing spots close together) are the
"solved" parts of the hole; rows with wide multi-colour scatter, or dots
sitting near the water/OB shading, are where the strategy is still genuinely
contested. This is the spatial complement to the `equiv_spatial_map.png` /
cross-seed Jaccard maps earlier in the report — those tell you *how much*
ambiguity there is per point, this shows *what it looks like on the ground*.

---

# Conclusions: why do some points never fully settle — even with equivalence sets, not a single arg-min?

The whole point of switching from "track the single best (club, aim)" to
"track the *set* of statistically tied-best options" was to stop penalising
the model for a problem that isn't really there: two combinations that score
the same shouldn't be treated as a disagreement. That mechanism works — a
genuine tie, once both members are in `E*`, has no reason to ever leave: if
two options are truly equal, there's no force pushing either one's estimate
away from the other, so a stable set should just keep recurring.

So a flat surface cannot, by itself, explain why 20/280 grid points still
never lock onto a stable `E*` by `N = 500` (mean cross-seed agreement
~0.51 Jaccard). Saying "the surface is flat" is necessary context, but it
is not yet a mechanism — the actual driver is more specific than that, and
it's worth being precise about it because the two things that follow
(flatness, and a shrinking threshold) are not independent, compounding
causes; they are one mechanism described from two ends.

**The real mechanism: hard-threshold membership testing does not converge for a true (or near-true) tie.**
`E*` is defined relative to `SE_min`, which *shrinks* as `N` grows. For an
option that is genuinely *worse* than the best, its estimated gap from
`R_min` grows relative to that shrinking `SE_min` as `N` increases — more
samples make a real difference more visible, so it exits `E*` decisively and
stays out. But for an option whose *true* value ties the best, there is no
such signal to sharpen: its estimated gap from `R_min` is driven entirely by
sampling noise of the same order as `SE_min` itself, at every `N`. The ratio
(estimated gap) / `SE_min` never concentrates — it just keeps re-randomising
as more samples come in, so the option drifts in and out of `E*`
indefinitely. This is not a flaw in the equivalence-set logic or the
simulation; it is a known, unresolved case in ranking-and-selection
statistics (the "indifference zone" problem: the sample size needed to
guarantee a stable selection literally diverges as the true gap between two
options goes to zero). A hard `Jaccard = 1`-for-3-steps test is, by
construction, exactly the kind of rule this hits.

Where does "the surface is flat" fit in, then? It's the reason so many
options are *close enough* to the best for this to bite in the first place.
From the equivalence-set anatomy earlier: even at points classed "ambiguous"
by cross-seed Jaccard, the *actual* ESHO spread inside `E*` tops out around
**0.02 strokes** — several adjacent aim values are, for practical purposes,
indistinguishable. And it's overwhelmingly an **aim-tie**, not a club-tie
(204/280 points classified "aim-driven" vs. 21 "club-driven" ambiguity) —
the model reliably identifies the right club; the 5 yd aim discretisation is
what leaves multiple near-identical options in contention, each one a
candidate for the flickering described above. Flatness sets up *how many*
options are vulnerable to this; it doesn't by itself cause any single one of
them to flicker.

**A second, genuinely distinct mechanism: rare, large-magnitude OB events add heavy-tailed noise exactly at that boundary.**
Confirmed directly (see the OB exposure diagnostic): at the specific 20
non-converged points, near-optimal combos do carry small (0.1-5%) real OB
probability, and the OB penalty is large relative to the otherwise tiny
residual noise there (`gap` of 0.4-1.2 strokes vs. `SE_min` of
0.003-0.015 strokes). A combination whose long-run mean should sit just
inside `E*` will occasionally draw one unlucky OB shot in a given `N`-step's
new sample and jump by roughly one `SE_min` — enough to (temporarily) exit
the set. This is a textbook symptom of estimating a mean under a
low-probability, high-magnitude mixture: the empirical mean's sampling
distribution is right-skewed, so it converges to "stable" more slowly than
the implicit Gaussian assumption behind `SE = sqrt(var/n)` expects, even
though nothing about the estimate is wrong.

**Net read:** none of this points to a defect in the OB logic, the
equivalence-set definition, or the simulation itself — it's two real,
identifiable sources of residual ambiguity: (1) a hard SE-relative
membership test that structurally cannot converge for a true or near-true
tie, made relevant at these particular points because the aim discretisation
leaves several options within an indifference zone of each other, and
(2) rare OB tail events colliding with an otherwise very low-noise estimate.
They show up at the same handful of long-range points because that's exactly
where the aim grid is coarsest relative to the angular precision needed
(feeding (1)) and where OB is reachable at all (feeding (2)). If reducing
the residual matters more than understanding it, the highest-leverage lever
is a finer aim grid (or reporting the *centroid* of `E*` rather than any
single member) — that shrinks the indifference zone itself, rather than
trying to out-sample or out-threshold a tie that, by construction, no amount
of additional simulation resolves.

---

# Correction: the OB fix did not change these results — the earlier motivating claim was based on a bad plot

The story used to introduce the OB logic elsewhere (including earlier
drafting of the paper's methods section) was: *shots were escaping past the
water hazard into an unpenalised "free" zone, artificially destabilising the
argmin at long-range grid points; adding the boundary + penalty closes that
gap and the numbers above are the corrected picture.* That causal claim
needs to be withdrawn.

The evidence for "shots were escaping unpenalised" came from a boundary
visualisation that was itself drawn incorrectly — a plotting error, not a
finding about the simulation. Once that is accounted for, there is no basis
left to claim the OB fix changed the convergence behaviour reported in this
document; the honest expectation is that the stabilisation and cross-seed
agreement numbers above are **the same, to within simulation noise, as they
would have been without the fix.** The two mechanisms explaining residual
non-convergence in the Conclusions section hold regardless — neither
depends on the OB fix having changed anything.

This does **not** mean the OB fix was pointless. It is still the correct
thing to have in the model: a shot travelling out of bounds should score a
stroke-and-distance penalty, full stop, independent of whether it happens to
move any convergence statistic. The fix should be kept and described in the
paper — just on its own merits (rule-correctness), not as an explanation for
the residual non-convergence documented above. The two-mechanism account
in the Conclusions section is the actual explanation, and it would apply
identically with or without OB in the model.

**What to change in the paper draft as a result:** the framing that
introduces OB (methods) should not claim it was "diagnosed as a source of
the non-convergence" or that the study was "re-run" to correct that specific
problem — that motivating narrative doesn't hold up. State plainly instead
that OB was added because it is a rules-correctness gap (a shot going out of
bounds must be penalised, and the pre-fix model didn't do that), independent
of the convergence analysis, and that the convergence numbers in Results
reflect the current, correct model rather than a "before vs. after OB"
comparison.


---

# In plain terms

Skip the statistics jargon above — here's the same finding in plain
language, with a picture.

Imagine two golfers' shots are genuinely, truly equally good — same expected
score, no difference at all. We want to know: does our test ever settle down
and say "yes, these two are tied, done"? You'd think the answer is
automatically yes. It isn't, and here's why.

Every time we simulate more shots, two things happen at once: our estimate
of each option's score gets a little more precise, **and** the "close enough
to count as tied" tolerance we use gets a little *stricter* (it has to — a
looser tolerance would call almost everything tied). For an option that's
truly *worse*, this is fine: more simulation makes the gap more obvious
faster than the tolerance tightens, so it clearly and permanently fails the
test. But for an option that's truly *tied*, there is no real gap for more
simulation to reveal — the tiny difference we measure between two truly
equal options is 100% random noise, and that noise shrinks at roughly the
same rate as the tolerance does. So the coin never stops flipping: at any
given amount of simulation, the tied option might happen to measure just
inside the tolerance, or just outside it, by pure chance — forever. That's
not our test failing; it's what happens whenever you ask a strict yes/no
question about two things that don't actually differ.

```{r plain-terms-diagram, fig.width=9, fig.height=5, echo=FALSE, warning=FALSE, message=FALSE}
set.seed(42)

N <- seq(10, 500, by = 10)
n_len <- length(N)
se_band <- 0.35 / sqrt(N)              # the shrinking "close enough" tolerance

# A truly tied option: pure noise on the same scale as the band, every step
tied_gap  <- rnorm(n_len, mean = 0, sd = se_band)
# A genuinely worse option: a small but real gap, buried in noise early on,
# that becomes obvious as the noise (not the true gap) shrinks
worse_gap <- 0.28 + rnorm(n_len, mean = 0, sd = se_band * 0.5)

df <- tibble(
  N = rep(N, 2),
  gap = c(tied_gap, worse_gap),
  band = rep(se_band, 2),
  option = rep(c("Truly tied with the best", "Genuinely a bit worse"), each = n_len)
)

ggplot(df, aes(N, gap)) +
  geom_ribbon(aes(ymin = -band, ymax = band), fill = "grey70", alpha = 0.35) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_line(aes(colour = option), linewidth = 0.9) +
  facet_wrap(~option, ncol = 1) +
  scale_colour_manual(values = c("Truly tied with the best" = "#0072B2",
                                  "Genuinely a bit worse" = "#D55E00"), guide = "none") +
  labs(
    title = "Why a true tie never settles, but a real gap does",
    subtitle = "Grey band = the shrinking 'close enough to count as tied' tolerance",
    x = "N (simulated shots)",
    y = "Estimated gap from the best option"
  ) +
  theme(strip.text = element_text(face = "bold"))
```

**Read the picture as:** the grey band is the tolerance, shrinking left to
right as we simulate more. The orange line (a real, small difference) starts
inside the band but climbs steadily out of it and stays out — more data
resolved a real question. The blue line (a true tie) has no such trend: it
just wanders randomly above and below zero, inside and outside the
shrinking band, no matter how far right you go. That's the entire
mechanism, visualised — and it's also why the honest fix is a *finer aim
grid* (fewer truly-tied options to begin with), not more simulation of the
same grid.

---

# Testing a new hypothesis: the fixed-yard aim grid's angular resolution shrinks with distance

**The diagnosis.** Every grid point tests the same fixed set of aim offsets
in *yards* (`-40, -35, ..., 35, 40`). But the simulation converts an aim
offset to an *angle* via `atan(aim_offset / total_dist)` — and for a fixed
yardage, that angle shrinks as `total_dist` grows. Concretely: at 40yd from
the pin, adjacent 5yd aim steps are ~7° apart; at 290yd, they're ~2° apart.
Far from the pin, the tested options are almost the same direction by
construction — no wonder they're hard to tell apart.

```{r aim-fan-setup}
pin_x <- 5; pin_y <- 334
aim_grid <- seq(-40, 40, by = 5)
stub_len <- 10

stability_pts <- gridpoints |> select(x, y) |> distinct() |>
  left_join(cross_at_max |> select(x, y, mean_jaccard), by = c("x", "y"))

fan_all <- stability_pts |>
  crossing(aim_offset = aim_grid) |>
  mutate(
    total_dist = sqrt((pin_x - x)^2 + (pin_y - y)^2),
    unit_x = (pin_x - x) / total_dist,
    unit_y = (pin_y - y) / total_dist,
    angle_rad = atan(aim_offset / total_dist),
    dir_x = cos(angle_rad) * unit_x - sin(angle_rad) * unit_y,
    dir_y = sin(angle_rad) * unit_x + cos(angle_rad) * unit_y,
    xend = x + dir_x * stub_len,
    yend = y + dir_y * stub_len
  )
```

## All tested aim directions, every grid point, coloured by stability

```{r aim-fan-all, fig.width=10, fig.height=11}
ggplot() +
  geom_segment(data = fan_all, aes(x = x, y = y, xend = xend, yend = yend, colour = mean_jaccard),
               linewidth = 0.35, alpha = 0.6) +
  geom_point(data = stability_pts, aes(x, y, colour = mean_jaccard), size = 0.8) +
  scale_colour_viridis_c(name = "cross-seed\nJaccard\n(stability)", limits = c(0, 1)) +
  coord_fixed() +
  labs(title = "Tested aim directions at every grid point (-40..+40y, 5y step), true bearing to pin",
       subtitle = "each stub = one tested aim_offset, fixed 10y length; colour = cross-seed agreement at N=500",
       x = "x (yards)", y = "y (yards)")
```

**What this shows.** Every one of the 280 grid points, with its full tested
aim grid drawn as a fan of short rays in the *true* direction (same
rotation as the corrected fan earlier — this reuses that fix), coloured by
how well the 100 seeds agree at that point.

**Read it as.** Near the top of the grid (close to the pin) the fans are
wide open umbrellas — the 17 tested directions are clearly, widely
separated — and they're mostly yellow/green (high agreement). Toward the
bottom (far from the pin) the fans collapse to something close to a single
line — the tested directions are nearly indistinguishable — and the colour
drops toward purple (low agreement). The visual pattern is the diagnosis.

## Quantifying it: angular fan width vs distance, and vs stability

```{r aim-fan-width, fig.width=8.5, fig.height=5.5}
fan_width <- fan_all |>
  filter(aim_offset %in% c(-40, 40)) |>
  select(x, y, total_dist, mean_jaccard, aim_offset, angle_rad) |>
  pivot_wider(names_from = aim_offset, values_from = angle_rad, names_prefix = "ang_") |>
  mutate(fan_width_deg = (`ang_40` - `ang_-40`) * 180 / pi)

ggplot(fan_width, aes(total_dist, fan_width_deg, colour = mean_jaccard)) +
  geom_point(size = 2.2) +
  scale_colour_viridis_c(name = "cross-seed\nJaccard", limits = c(0, 1)) +
  labs(title = "Angular resolution of the fixed-yard aim grid vs distance to pin",
       subtitle = "fan_width_deg = angle between the aim=-40y and aim=+40y rays; colour = stability",
       x = "distance from grid point to pin (yards)", y = "angular fan width (degrees)")
```

```{r aim-fan-cor}
tibble(
  comparison = c("fan width vs distance", "fan width vs stability (Jaccard)",
                 "distance vs stability (Jaccard)"),
  correlation = c(
    cor(fan_width$fan_width_deg, fan_width$total_dist),
    cor(fan_width$fan_width_deg, fan_width$mean_jaccard),
    cor(fan_width$total_dist, fan_width$mean_jaccard)
  )
) |> kable(digits = 3, caption = "Pearson correlation across the 280 grid points")
```

**What this shows.** The top-left curve is a pure geometric fact — no
noise, it's just `atan()` — the angular gap between the extreme tested aim
offsets falls from ~78° near the pin to ~16° at the far end of the grid. The
colour on that same plot is the empirical stability at each point, and the
correlation table quantifies both relationships at once.

**Read it as.** The diagnosis holds up, honestly stated: fan width vs
stability correlates at **r ≈ 0.64** — a real, moderately strong
relationship, visible directly in the first plot as the colour gradient
running top-to-bottom with the fan width. It is not a perfect 1-to-1
explanation (r would be ~1 if it were) — points at similar distances still
show some colour scatter, meaning other point-specific factors (hazard
proximity, OB exposure) also contribute, consistent with the two-mechanism
account earlier in this document. But the aim grid's shrinking angular
resolution is a real, substantial, and previously undocumented contributor
to why long-range points are harder to pin down — a finer aim grid
specifically at longer range (rather than uniformly across all 280 points)
would be the most targeted fix.

## Proposed fix: a fixed-angle aim grid instead of a fixed-yard one

**The change.** Replace the fixed-yard `aim_offset` grid with a fixed-*angle*
grid, and derive the yard offset per point instead of the other way round:

$$\theta \in \{-40°, -35°, \dots, 35°, 40°\} \qquad \text{aim\_offset}(d) = d \cdot \tan(\theta)$$

**This is what's now actually implemented** in `core.py::simulate_approach_shots`
(the `aim_angle_range`/`aim_angle_step` parameters, on by default in
`convergence_worker.py`/`run_hpc_worker.py` — see `--no-aim-angle` to fall
back to the old fixed-yard grid). 17 angles spanning $\pm 40°$ in $5°$
steps — the same option count as the original fixed-yard grid, but now with
constant angular resolution at every distance instead of resolution that
shrinks far from the pin: two earlier, more aggressively pruned drafts
(±16°/2°, then ±30°/5°, then ±30°/10°) were considered and rejected in
favour of this one, which keeps near-green steps close to the original
~5-yard granularity the yard grid gave there, while still fixing the
far-range crowding. Mechanically this is still a
*simplification*,
not an added step: `core.py`'s `rotation_translator` already consumes an
angle, not a yard offset — the yard offset only ever existed so that
`angle_deg = atan(aim_offset / total_dist)` could produce one. A fixed-angle
grid skips that conversion entirely; if a yard label is still wanted for
reporting, it is now the *derived* quantity, computed once per point from
the fixed angle.

```{r fixed-angle-setup}
theta_deg <- seq(-40, 40, by = 5)   # matches core.py's default: 17 angles, +-40 deg, 5 deg step
theta_rad <- theta_deg * pi / 180

fan_fixed_angle <- stability_pts |>
  crossing(theta_deg = theta_deg) |>
  mutate(
    total_dist = sqrt((pin_x - x)^2 + (pin_y - y)^2),
    theta_rad  = theta_deg * pi / 180,
    aim_offset_yd = total_dist * tan(theta_rad),  # derived, not fixed
    unit_x = (pin_x - x) / total_dist,
    unit_y = (pin_y - y) / total_dist,
    dir_x  = cos(theta_rad) * unit_x - sin(theta_rad) * unit_y,
    dir_y  = sin(theta_rad) * unit_x + cos(theta_rad) * unit_y,
    xend = x + dir_x * stub_len,
    yend = y + dir_y * stub_len
  )

# per-point labels: the max (=|min|, symmetric) yard offset actually tested,
# and what one 5 deg step corresponds to in yards at that point's distance
point_labels <- stability_pts |>
  mutate(
    total_dist   = sqrt((pin_x - x)^2 + (pin_y - y)^2),
    max_offset_yd  = total_dist * tan(40 * pi / 180),
    step_offset_yd = total_dist * tan(5 * pi / 180),
    label = sprintf("+-%.0fy\nstep %.1fy", max_offset_yd, step_offset_yd)
  )
```

```{r fixed-angle-plot, fig.width=11, fig.height=13}
ggplot() +
  geom_segment(data = fan_fixed_angle, aes(x = x, y = y, xend = xend, yend = yend, colour = mean_jaccard),
               linewidth = 0.45, alpha = 0.6) +
  geom_point(data = stability_pts, aes(x, y, colour = mean_jaccard), size = 0.8) +
  geom_text(data = point_labels, aes(x, y, label = label), size = 1.55, vjust = 2.3,
            lineheight = 0.85, colour = "grey20") +
  scale_colour_viridis_c(name = "cross-seed\nJaccard\n(stability)", limits = c(0, 1)) +
  coord_fixed() +
  labs(title = "As implemented: fixed +-40 deg / 5 deg-step aim grid (17 clearly distinct options everywhere)",
       subtitle = "fan width is now angle-constant by construction; label at each point = max yd offset tested (+-) and yd/step at that distance",
       x = "x (yards)", y = "y (yards)")
```

**Read it as.** Directly underneath the diagnostic plot above: same points,
same colour scale, same stub length — but every fan is now the same
*angular* width, keeping the same 17-option count as the original fixed-yard
grid instead of the more aggressively pruned 7-option ±30°/10° draft
considered above. Near the pin the window is ~±42y with a ~4.4y step
(close to the original fixed grid's own ~5-yard granularity there, so
near-green resolution barely changes); far from the pin the window opens
out to ~±243y with a ~25.3y step — considerably wider than the old fixed
±40y, on purpose: several of those extreme far-range options will be
obviously bad (off the fairway, OB) and should exit the equivalence set
decisively rather than hover near it, which is exactly the well-behaved
case in the "why a true tie never settles" mechanism above.

```{r fixed-angle-range-table}
point_labels |>
  summarise(
    `min distance (yd)` = min(total_dist), `max distance (yd)` = max(total_dist),
    `max offset at min dist (yd)` = min(max_offset_yd), `max offset at max dist (yd)` = max(max_offset_yd),
    `step at min dist (yd)` = min(step_offset_yd), `step at max dist (yd)` = max(step_offset_yd)
  ) |>
  pivot_longer(everything(), names_to = "quantity", values_to = "value") |>
  kable(digits = 1, caption = "Range of implied yard offsets/steps across the 280-point grid under the fixed +-40 deg / 5 deg grid now implemented in core.py")
```
